The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-02-17 src/ms.tex
capitalize SkyCoord shortcut
Signed-off-by: nstarman <nstarkman@protonmail.com>

\subsubsection*{New Features} \label{sec:core-features-new}

\begin{itemize}
%  \item Time Series sub-package (V3.1: New Box Least Squares Periodogram)
%  \item Uncertainties and Distributions
  \item ERFA module is now it’s own thing; Obliquely tie in that units
    remains in astropy
  \item ? Pytest plugin ecosystem
  \item A new SpectralCoord class for representing and transforming spectral quantities
    Velocity in Coordinates
  \item WCS API : A shared Python interface for World Coordinate Systems
        Include APE reference
\end{itemize}

New features come in various forms, from new submodules providing
wholly new capabilities, to larger additions to existing submodules,
to large rewrites of the interface of modules.  This range is spanned
by the four main features that entered the astropy core, which we
describe below.

@@

This \astropyDistribution\ can be any type of array, including of
course a \astropyQuantity, and the goal is to ensure it can be
seamlessly used to instantiate also other astropy classes, such as
\astropyskycoord\ and \astropyTime.

@@

>>>>

\subsubsection*{New and Planned Features} \label{sec:core-features-new}

% \begin{itemize}
%  \item Time Series sub-package (V3.1: New Box Least Squares Periodogram)
%  \item Uncertainties and Distributions
%   \item ERFA module is now it’s own thing; Obliquely tie in that units
%    remains in astropy -> above
%  \item ? Pytest plugin ecosystem -> done elsewhere
%  \item A new SpectralCoord class for representing and transforming spectral quantities
%    Velocity in Coordinates
%  \item WCS API : A shared Python interface for World Coordinate Systems
%        Include APE 14 reference
% \end{itemize}

New features come in various forms, from new submodules providing
wholly new capabilities, to larger additions to existing submodules,
to large rewrites of the interface of modules.  This range is spanned
by the five main new features that entered the astropy core since the
previous summary \citepalias{astropy:2018}.

New features present a particular difficulty for a mature package such
as \astropypkg, because by their nature they have not been tried out
much and it is difficult to know in advance what the best interface
will be, and hard to address all possible use cases in one go, even
within \astropypkg\ itself.  To give a sense of this, we include in
our descriptions how we hope the new features will evolve (realizing
that, in the end, in a user-driven project like \astropy, what happens
depends mostly on what is highest on the priority list of those that
have time to contribute).

@@

This \astropyDistribution\ can be any type of array, including of
course a \astropyQuantity, and the goal is to ensure it can be
seamlessly used to instantiate also other astropy classes, such as
\astropySkyCoord\ and \astropyTime.

Apart from Monte-Carlo error propagation, the plan for
\astropysubpkg{uncertainties} is to also support the case where values
have associated uncertainties, and these are propagated assuming that
they are normally distributed, but tracking covariances.  One
implementation problem to be solved is when to {\em stop} tracking
covariances.  For instance, a simple operation such as subtracting the
mean from $N$ data points implies that all data points are now
covariant with each other, i.e., one has to carry $N\times N$
covariances.  For a large image, that becomes not just pointless but
also prohibitive in terms of memory and CPU usage.

@@

\paragraph{Spectral Coordinates.} Measurements are often taken at
specific ``spectral coordinates'', be they frequencies, wavelengths,
or photon energies.  The new \astropySpectralCoord\ class allows to
transform between these as well as to equivalent Doppler velocities.
It builds on the \astropyQuantity, which can also represent these
kinds of physical quantities, and allows conversion via dedicated
equivalencies, but aims to make this more straightforward.
Furthermore, it can be made aware of the observer and target reference
frames, allowing for instance transformation from telescope-centric (or
topocentric) frames to Barycentric or Local Standard of Rest
(LSRK and LSRD) velocity frames.

\paragraph{A High-Level Interface to World Coordinate Systems.}
Astronomical data are often provided with information about the
``real-world'' coordinates that correspond to pixel coordinates. This
mapping is the essence of the ``World Coordinate System (WCS)''
concept.  From its inception, the \astropysubpkg{wcs} sub-package
allowed access to WCS information provided in, e.g., FITS files, but
it became clear that other WCS standards and representations had to be
supported for new missions and observatories (e.g., the James Webb
Space Telescope and the Rubin Observatory).  To harmonize these, a new
high-level interface was created, based on a formal design first
proposed via an \astropy\ Proposal for Enhancement (\astropyAPE{14}).
The hope is that by having a formal design, it will be straightforward
for other packages implementing WCS objects to either modify their
classes to conform to the new interface or build thin wrappers that
conform to it.  The implementation in \astropysubpkg{wcs} interacts
well with other \astropypkg\ objects such as \astropySkyCoord\ and
\astropyTime.
